---
title: "Chapter 2 - Data Visualization"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
#knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```


## Scatterplots

```{r, message=FALSE}
library(nycflights13)
library(dplyr)
library(ggplot2)
```

### Alaska Airlines Inc. flights only

```{r}
alaska_flights <- flights %>% 
  filter(carrier == 'AS')
```

### Standard Scatterplot

```{r}
ggplot(data = alaska_flights, mapping = aes(x = dep_delay, y = arr_delay)) +
  geom_point()
```


### Prevent Overplotting: Adjust Transparency of Points

```{r}
ggplot(data = alaska_flights, mapping = aes(x = dep_delay, y = arr_delay)) +
  geom_point(alpha = 0.2)
```

### Prevent Overplotting: Jittered Points

```{r}
ggplot(data = alaska_flights, mapping = aes(x = dep_delay, y = arr_delay)) +
  geom_jitter(width = 30, height = 30)
```


## Linegraphs


```{r}
glimpse(weather)
```

```{r}
early_january_weather <- weather %>% 
  filter(origin == "EWR", month == 1, day <= 15)
```

### Newark Temperature (C) in early January 2013

```{r}
ggplot(data = early_january_weather, aes(x = time_hour, y = (temp - 32) * 5/9)) +
  geom_line()
```

### Newark Temperature (C) in 2013

```{r}
jan_feb_weather <- weather %>% 
  filter(origin == "EWR", month < 3) %>% 
  mutate(temp = (temp - 32) * 5/9)

ggplot(jan_feb_weather, aes(x = time_hour, y = temp)) +
  geom_line()
```


## Histograms

```{r}
ggplot(data = weather, mapping = aes(x = temp)) +
  geom_histogram()
```

```{r}
ggplot(data = weather, mapping = aes(x = temp)) +
  geom_histogram(color = "white", breaks = seq(0, 100, 10))
```

## Facets

```{r}
ggplot(data = weather, mapping = aes(x = temp)) +
  geom_histogram(binwidth = 5, color = "white") +
  facet_wrap(~ month)
```

```{r}
month_names = as_labeller(
  c(`1` = "January", `2` = "February", `3` = "March", `4` = "April",
    `5` = "May", `6` = "June", `7` = "July", `8` = "August",
    `9` = "September", `10` = "October", `11` = "November", `12` = "December"
    )
  )

ggplot(data = weather, mapping = aes(x = temp)) +
  geom_histogram(binwidth = 5, color = "white") +
  facet_wrap(~ month, nrow = 4, labeller = month_names)
```


## Boxplots

```{r}
ggplot(data = weather, mapping = aes(x = factor(month), y = temp)) +
  geom_boxplot()
```

Which months have the highest variability in temperature?

```{r}
iqr <- weather %>% 
  group_by(month) %>% 
  summarise(IQR = IQR(temp, na.rm = TRUE)) %>% 
  arrange(desc(IQR))

knitr::kable(iqr)
```


## Barplots

Which geom to use depends on how your data is represented.

#### Raw data (not pre-counted): use `geom_bar()`

Like categorical data in a column of a data frame.

```{r}
fruits <- tibble(
  fruit = c("apple", "apple", "orange", "apple", "orange")
)
knitr::kable(fruits)
```

```{r}
ggplot(data = fruits, mapping = aes(x = fruit)) +
  geom_bar()
```


#### Pre-counted data: use `geom_col()`

Here the fruit category and its count are in different columns.

```{r}
fruits_counted <- tibble(
  fruit = c("apple", "orange"),
  n = c(3, 2)
)
knitr::kable(fruits_counted)
```

Use `geom_col()` with the y-position aesthetic mapped to the variable that has
the counts.

```{r}
ggplot(data = fruits_counted, mapping = aes(x = fruit, y = n)) +
  geom_col()

```

#### Using `geom_bar()` to visualize the occurance of carriers

```{r}
ggplot(data = flights, mapping = aes(x = carrier)) +
  geom_bar()
```


#### To use `geom_col()` the occurance of each carrier first has to be counted

```{r}
carriers_counted <- flights %>%
  count(carrier)
knitr::kable(carriers_counted)
```


```{r}
ggplot(data = carriers_counted, mapping = aes(x = carrier, y = n)) +
  geom_col()
```


## Two categorical Variables

An additional categorical variable can be displayed as fill color.

Since this is also an aesthetic mapping it has to be placed inside `aes()`.

```{r}
ggplot(data = flights, mapping = aes(x = carrier, fill = origin)) +
  geom_bar()
```

**CAVE:** the `color` parameter only changes the outline color of bars.

```{r}
ggplot(data = flights, mapping = aes(x = carrier, color = origin)) +
  geom_bar()
```


## Side-by-side or *dodged* barplots

Default barplots are `stacked`. Which makes it sometimes difficult compare
values (e.g. for each `origin`).  An alternative are `dodged` barplots.

```{r}
ggplot(data = flights, mapping = aes(x = carrier, fill = origin)) +
  geom_bar(position = "dodge")
```

To make the bars same width, even if a carrier has only one or two origin
airports we can use the more robust `position_dodge()` function.

```{r}
ggplot(data = flights, mapping = aes(x = carrier, fill = origin)) +
  geom_bar(position = position_dodge(preserve = "single"))
```


## Faceted Barplots

```{r}
ggplot(data = flights, mapping = aes(x = carrier)) +
  geom_bar() +
  facet_wrap(~ origin, ncol = 1)
```



## Pie Chart

Usage is not recommended, since it is not easy to see differences.

```{r}
ggplot(data = carriers_counted, mapping = aes(x = "", y = n, fill = carrier)) +
  geom_bar(stat="identity", width=1, color = "white") +
  coord_polar("y", start=0) +
  theme_void()
```

