---
title: "Chapter 3 - Data Wrangling"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
```

## Important functions of the `dplyr` Package

Most important `dplyr` functions are:

* `filter()` - pick only a subset of the data (SQL **where clause**).
* `summarize()` - run *summary statistics*.
* `group_by()` - create sub-groups of your data; allowing summary statistics
for each group.
* `mutate()` - add variables/columns to the data frame.
* `arrange()` - change order of the records (SQL **order by**)
* `join()` - combine one data frame with another.


```{r, message=FALSE}
library(dplyr)
library(ggplot2)
library(nycflights13)
```

## Subset Observations / Rows with `filter()`

```{r}
portland_flights <- flights %>% 
  filter(dest == "PDX")

glimpse(portland_flights)
```

Filter flights for all rows that departed from JFK and were heading to Burlington, Vermont ("BTV") or Seattle, Washington ("SEA") and departed in the months of October, November, or December.

```{r}
btv_sea_flights_fall <- flights %>% 
  filter(origin == "JFK" & (dest == "BTV" | dest == "SEA") & month > 9)

dim(btv_sea_flights_fall)
```

Criteria can also be separated by commas:

```{r}
btv_sea_flights_fall <- flights %>% 
  filter(origin == "JFK", (dest == "BTV" | dest == "SEA"), month > 9)

dim(btv_sea_flights_fall)
```

### Use `%in% c()` for comparison with a vector

```{r}
many_airports <- flights %>% 
  filter(dest %in% c("SEA", "SFO", "PDX", "BTV", "BDL"))

dim(many_airports)
```


## `summarize()` Variables

Compute *summary statistics* like:

* `mean()`: the average
* `sd()`: the standard deviation, which is a measure of spread
* `min()` and `max()`: the minimum and maximum values, respectively
* `IQR()`: interquartile range
* `sum()`: the total amount when adding multiple numbers
* `n()`: a count of the number of rows in each group

```{r}
summary_temp <- weather %>% 
  summarize(mean = mean(temp), std_dev = sd(temp))

knitr::kable(summary_temp)
```

Let's ignore missing values:

```{r}
summary_temp <- weather %>% 
  summarize(
    mean = mean(temp, na.rm = TRUE),
    std_dev = sd(temp, na.rm = TRUE),
    count = n())

knitr::kable(summary_temp)
```


## `group_by()` rows

Summarize temperature for every month:

```{r}
summary_monthly_temp <- weather %>% 
  group_by(month) %>% 
  summarize(    
    mean = mean(temp, na.rm = TRUE),
    std_dev = sd(temp, na.rm = TRUE),
    count = n()) %>% 
  mutate(month = month.name[month])

knitr::kable(summary_monthly_temp)
```

Grouping does not change the data, but the meta data.
Here diamonds grouped by *cut*:

```{r}
diamonds %>% 
  group_by(cut)
```

